---
title: "Global shortfalls in documented actions to conserve biodiversity"
output:
    bookdown::word_document2:
        reference_docx: ms_general.docx
        number_sections: false
    bookdown::pdf_document2:
        toc: no
        latex_engine: xelatex
        fig_caption: yes
        keep_tex: yes
link-citations: yes
csl: nature.csl
biblio-style: apalike
bibliography: [refs.bib, funky-refs.bib, pckgs.bib]
fontsize: 12pt
geometry: margin=2.5cm
mainfont: Calibri
lang: en-GB
papersize: a4
documentclass: article
header-includes:
- \usepackage{hyperref}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- \fancyfoot[CO,CE]{\thepage}
- \usepackage{setspace}
- \doublespacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,error = FALSE, warning = FALSE,
                      dpi = 300, fig.path = "figs/main/", fig.align = "center", 
                      fig.width = 16.6/2.54, fig.height = 15/2.54)
```

```{r packages}
# Package names
packages <-  c(
    "dplyr", "tidyr","purrr", # Data wrangling
    "sf", "png", # Loading data
    "ggplot2", "ggtext", "ggnewscale","grid","gridExtra","scales",
    "RColorBrewer","cowplot","egg","rphylopic",
    "DiagrammeR","DiagrammeRsvg","rsvg")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
    install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

```{r functions}
myTheme <- function(base_size = 12, base_family = "") 
{
  theme_bw(base_size = base_size, base_family = base_family) %+replace% 
    theme(axis.text = element_text(size = 8, 
                                   colour = "black"),
          axis.title = element_text(size = 10, 
                                    vjust = 1, 
                                    face = "bold"),
          legend.text = element_text(size = 8, 
                                   colour = "black"),
          legend.title = element_text(size = 10),
          strip.text = element_text(size = 10,
                                    margin = margin(0.1,0.1,0.1,0.1, "cm")),
          strip.background = element_blank(),
          panel.grid = element_blank())
}

# Define function to capitalise first letter
capitalise <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}

# Function to make significance levels pretty
siglevel <- function(pval){
  ifelse(pval < 0.001, "P < 0.001",
         ifelse(pval < 0.05, 
                paste("P = ", signif(pval, 2), sep = ""),
                "P > 0.05 ")
         )
}

# Function factory for getting integer y-axis values.
integer_breaks <- function(n = 5, ...) {
  fxn <- function(x) {
    breaks <- floor(pretty(x, n, ...))
    names(breaks) <- attr(breaks, "labels")
    breaks
  }
  return(fxn)
}

# Function to calculate inverse logit
inv.logit <- function(x) exp(x)/(1 + exp(x))

# Function to calculate critical significance value
ci.critical <- function(siglevel) {
  qnorm((100 - siglevel) / 100 / 2, lower.tail = FALSE)
}
```

```{r data, include = FALSE}
# Main data
alliucn <- readRDS("data/analysis/main/all_iucn_prepped.Rds") 
```

```{r plotting-params}
imgs <- readRDS("figs/phylopic/phylopics.Rds")

# Useful lookups
class_lookup <- c("Malacostraca","Amphibia", "Aves", "Mammalia", "Reptilia",
                  "Cycadopsida", "Magnoliopsida", "Pinopsida")
names(class_lookup) <- tolower(class_lookup)
threat_lookup <- c(
  threat_habitat_change = "Habitat change",
  threat_overexploitation = "Overexploitation",
  threat_inv_spp = "Invasive species"
)

# Palettes
propPal <- colorRampPalette(brewer.pal(9, "Greens"))(15)
iucnPal <- RColorBrewer::brewer.pal(9, "YlOrRd")[c(4,6,8)]
cbpal <- RColorBrewer::brewer.pal(11, "RdBu")

# Other
matched_threats <- c("threat_habitat_change", "threat_resource_use", "threat_inv_spp" )
matched_actions <- c("matched_hc", "matched_ru", "matched_is")
ci_const <- ci.critical(95)
paProps <- c(1,5,10,25,50,75,90,95,99)
```

```{r sort-data}
# Spatial endemics -------------------------------------------------------------
endemic_realm <-
  alliucn %>% 
  filter(
    !(is.na(realms)),
    !(grepl(";", realms))) %>% 
  dplyr::select(scientificName, className, realms, actions_zero_sps) %>% 
  separate_rows(realms, sep = ";") %>% 
  rename(realm = realms)
endemic_country <-
  alliucn %>% 
  filter(
    !(is.na(countries)),
    !(grepl(";", countries))) %>% 
  dplyr::select(scientificName, className, countries, actions_zero_sps) %>% 
  separate_rows(countries, sep = ";") %>% 
  rename(country = countries)

# Sum threatened endemic species by country
ctry_threatspp <-
  group_by(endemic_country, country) %>% 
  summarise(n = n())

# Threats unknown --------------------------------------------------------------
threat_unknown_count <- readRDS("data/threat_unknown_count.Rds")

# Threat presence/absence ------------------------------------------------------
alliucn_threats <-
  gather(alliucn, key = "threat", value = "threat_present", 
         starts_with("threat_"), -"threat_count") %>% 
    # Relevel
    mutate(threat = factor(threat,
                           levels = c("threat_habitat_change",
                                      "threat_resource_use", 
                                      "threat_inv_spp", 
                                      "threat_pollution",
                                      "threat_climate_change",
                                      "threat_other")),
           className =  factor(className,
                               levels = sort(unique(className)))) %>% 
  # Drop NAs
  filter(!(is.na(threat_present))) %>% 
  dplyr::select(scientificName, className, iucn, ED, EDGE.Rank,cat_change_bi,
                threat, threat_present, threat_count, 
                starts_with("matched_"), 
                starts_with("action_count_"))

# EDGE -------------------------------------------------------------------------
alliucn_EDGE <-
  alliucn %>% 
  filter(className %in% c("amphibia","aves","mammalia"),
         !(is.na(ED))) %>%  
  mutate(EDGE = case_when(is.na(EDGE.Rank) ~ FALSE,
                          !(is.na(EDGE.Rank)) ~ TRUE),
         ED = as.numeric(ED)) %>% 
  dplyr::select(scientificName, className, iucn, EDGE, ED, actions_zero_sps) %>% 
  # Standardise ED within taxonomic class to make comparable between classes
  group_by(className) %>% 
  summarise(ED_max = max(ED),
            ED_min = min(ED),
            ED_stand = ED / max(ED),
            scientificName = scientificName, 
            iucn = iucn, 
            EDGE = EDGE, 
            ED = ED, 
            actions_zero_sps = actions_zero_sps)

# Matched action presence/absence -----------------------------------------------
alliucn_matched <- 
  alliucn_threats %>% 
  filter(
    # Threat is present
    threat_present == 1,
    # Threat is  habitat change/resource use/invasive species
    threat %in% matched_threats
    ) %>% 
  # Relevel
  mutate(threat = factor(threat,
                         levels = c("threat_habitat_change",
                                    "threat_resource_use",
                                    "threat_inv_spp"),
                         labels = c("threat_habitat_change",
                                    "threat_overexploitation",
                                    "threat_inv_spp")))

# Status change ----------------------------------------------------------------
statchange <- readRDS("data/analysis/main/status_change_prepped.Rds")
statchange_matched <-
  # Gather threats
  gather(statchange, key = "threat", value = "threat_present", 
         starts_with("threat_"), -"threat_count") %>% 
  filter(
    # Only considering habitat change/resource use/invasive species
    threat %in% matched_threats,
    # Threat is present
    threat_present == 1,
    # Only one major threat
    threat_count <= 1
    ) %>% 
  # Relevel
  mutate(rangeArTot = as.numeric(rangeArTot),
         LogRangeArea = log(rangeArTot)) %>% 
  dplyr::select(scientificName, className, familyName, 
                threat, threat_count, 
                cat_old_cat, change_reason, cat_change_bi,
                LogBodyMass, iucn, islandSp_95, 
                starts_with("matched_action_"), starts_with("action_count_"))
```

```{r action-zero-raw}
# Sort data
actions_zero_props <- names(alliucn)[grep("actions_zero_",names(alliucn))]

zero_action_sum <- 
  lapply(actions_zero_props, function(prop){
    alliucn %>% 
          group_by(className, iucn) %>% 
          summarise(N = length(which(!(is.na(!!sym(prop))))),
                    nT = sum(!!sym(prop), na.rm = TRUE)) %>% 
          filter(N != 0) %>% 
          mutate(className = droplevels(factor(className,
                                    levels = names(class_lookup),
                                    labels = class_lookup)),
                 classNumeric = as.numeric(className),
                 paProp = prop)
  })
zero_action_sum <-
  lapply(zero_action_sum, function(i){
    i %>% 
      left_join(group_by(i, className) %>% 
                  summarise(classN = sum(N)),
                by = "className")
  })
names(zero_action_sum) <- actions_zero_props
zero_action_sum <- zero_action_sum$actions_zero_sps

# Total across threat status
zero_action_class <-
  zero_action_sum %>% 
  group_by(className) %>% 
  summarise(nT = sum(nT),
            N = sum(N),
            prop = 100 * (nT / N))
```

```{r action-matched-raw}
matched_actions_all <- c("matched_hc_sps","matched_ru", "matched_is")

# Sort data
matched_action_sum <- 
    lapply(matched_actions_all, function(i){
        # Identify threat from matched action i
        threat <- ifelse(grepl("_hc_",i), "threat_habitat_change",
                         ifelse(grepl("_ru",i), "threat_overexploitation",
                                ifelse(grepl("_is",i), "threat_inv_spp",
                                       NA)))
        result <- 
            # Filter to this threat only
            alliucn_matched[alliucn_matched$threat == threat,] %>%
            # Group by taxonomic class & RL category 
            group_by(className, iucn, .drop = FALSE) %>% 
            summarise(
                # Count all species considered (i.e. not NA)
                N = length(which(!(is.na(!!sym(i))))),
                # Count species with matched action
                nT = sum(!!sym(i), na.rm = TRUE)
            ) %>%
            ungroup() %>% 
            complete(className,
                     iucn,
                     fill = list(N = NA, nT = NA)) %>%
            filter(!(is.na(iucn))) %>% 
            mutate(threat = threat,
                   matched_action = i,
                   className = factor(className,
                                      levels = names(class_lookup),
                                      labels = class_lookup),
                   classNumeric = as.numeric(className)
            )
        return(result)
    }) %>% 
    bind_rows()

# How many species assessed in each class and threat
matched_action_sum_sum <-
    matched_action_sum %>% 
    group_by(threat,matched_action,className) %>% 
    summarise(classN = sum(N, na.rm = TRUE))

# Join
matched_action_sum <- 
    left_join(matched_action_sum, matched_action_sum_sum, 
              by = c("threat","matched_action","className")) %>% 
    mutate(
        threat = factor(threat,
                        levels = c("threat_habitat_change",
                                   "threat_overexploitation",
                                   "threat_inv_spp"),
                        labels = threat_lookup),
        matched_action = factor(matched_action,
                                levels = c("matched_sps",
                                           "matched_hc_target", "matched_hc_targetRL",
                                           paste("matched_hc_",paProps, sep = ""),
                                           "matched_ru","matched_is"))
    )
rm(matched_action_sum_sum)

# Total across threat status
matched_action_class <-
  matched_action_sum %>% 
  group_by(threat, className) %>% 
  summarise(nT = sum(nT),
            N = sum(N),
            prop = 100 * (nT / N))

# Total across class and threat status
matched_action_total <-
  matched_action_sum %>% 
  group_by(threat) %>% 
  summarise(nT = sum(nT, na.rm = TRUE),
            N = sum(N, na.rm = TRUE),
            prop = 100 * (nT / N))
```

```{r action-matched-raw-RL}
matched_action_hc_targetRL_sum <- 
    # Filter to this threat only
    alliucn_matched[alliucn_matched$threat == "threat_habitat_change",] %>%
    # Group by taxonomic class & RL category 
    group_by(className, .drop = FALSE) %>% 
    summarise(
        # Count all species considered (i.e. not NA)
        N = length(which(!(is.na(matched_hc_targetRL)))),
        # Count species with matched action
        nT = sum(matched_hc_targetRL, na.rm = TRUE),
        prop = 100 * (nT / N)
    ) %>%
    ungroup() %>% 
    complete(className,
             fill = list(N = NA, nT = NA)) %>%
        mutate(threat = "threat_habitat_change",
           matched_action = "matched_hc_targetRL",
           className = factor(className,
                              levels = names(class_lookup),
                              labels = class_lookup),
           classNumeric = as.numeric(className)
    )

# Total across class and threat status
matched_action_hc_targetRL <-
  matched_action_hc_targetRL_sum %>% 
  summarise(nT = sum(nT, na.rm = TRUE),
            N = sum(N, na.rm = TRUE),
            prop = 100 * (nT / N))
```

```{r inv-spp}
invspp <- read.csv("data/invasiveSp.csv")
invspp_neglected <- 
  alliucn[which(alliucn$threat_inv_spp & alliucn$invasive_ctrl_success == 0),]
# What invasive species are threatening amphibs that lack control? 
invspp_neglect_sum_amphib <-
  table(invspp$invasiveSp[
    which(gsub(" ", "_", invspp$scientificName) %in% 
            invspp_neglected$scientificName[invspp_neglected$className == "amphibia"])])
invspp_neglect_sum_mamm <-
  table(invspp$invasiveSp[
    which(gsub(" ", "_", invspp$scientificName) %in% 
            invspp_neglected$scientificName[invspp_neglected$className == "mammalia"])])
```

```{r model-results}
# Results
load("models/main/results_match.Rdata")
load("models/main/results_zero.Rdata")
load("models/main/results_zero_country.Rdata")
load("models/main/results_zeroProb_country.Rdata")
load("models/main/results_actionZero_EDGE.Rdata")
```

```{r statchange-sum}
# All status changes
statchange_sum <-
  statchange %>% 
  mutate(cat_change_bi = factor(cat_change_bi)) %>% 
  group_by(className, cat_change_bi, .drop = FALSE) %>% 
  summarise(n = n())
statchangespp <- unique(statchange$scientificName)

# HC
statchange_match_hc_sum <-
  alliucn_matched %>% 
  filter(!(is.na(cat_change_bi)),
         scientificName %in% statchangespp,
         threat == "threat_habitat_change") %>% 
  mutate(matched_hc_sps = factor(matched_hc_sps)) %>% 
  group_by(className, matched_hc_sps, .drop = FALSE) %>% 
 summarise(n_improve = length(which(cat_change_bi == 1)),
            n_decline = length(which(cat_change_bi == 0)),
            n = n(),
            prop_decline = n_decline / n,
           ratio_improve = n_improve / n_decline) %>% 
  filter(className %in% c("amphibia","aves","magnoliopsida","mammalia","pinopsida")) %>% 
  mutate(matched_hc_sps = factor(matched_hc_sps, levels = c(1,0)),
         matched_action = "threat_habitat_change") %>% 
  rename("value" = "matched_hc_sps")

# RU
statchange_match_ru_sum <-
  alliucn_matched %>% 
  filter(!(is.na(cat_change_bi)),
         scientificName %in% statchangespp,
         threat == "threat_overexploitation") %>% 
  mutate(matched_ru = factor(matched_ru)) %>% 
  group_by(className, matched_ru, .drop = FALSE) %>% 
 summarise(n_improve = length(which(cat_change_bi == 1)),
            n_decline = length(which(cat_change_bi == 0)),
            n = n(),
            prop_decline = n_decline / n,
           ratio_improve = n_improve / n_decline) %>% 
  filter(className %in% c("amphibia","aves","magnoliopsida","mammalia","pinopsida")) %>% 
  mutate(matched_ru = factor(matched_ru, levels = c(1,0)),
         matched_action = "threat_overexploitation") %>% 
  rename("value" = "matched_ru")

# IS
statchange_match_is_sum <-
  alliucn_matched %>% 
  filter(!(is.na(cat_change_bi)),
         scientificName %in% statchangespp,
         threat == "threat_inv_spp") %>% 
  mutate(matched_is = factor(matched_is)) %>% 
  group_by(className, matched_is, .drop = FALSE) %>% 
 summarise(n_improve = length(which(cat_change_bi == 1)),
            n_decline = length(which(cat_change_bi == 0)),
            n = n(),
            prop_decline = n_decline / n,
           ratio_improve = n_improve / n_decline) %>% 
  filter(className %in% c("amphibia","aves","magnoliopsida","mammalia","pinopsida")) %>% 
  mutate(matched_is = factor(matched_is, levels = c(1,0)),
         matched_action = "threat_inv_spp") %>% 
  rename("value" = "matched_is")

# Bind
statchange_match_sum <- 
  rbind(statchange_match_hc_sum, statchange_match_ru_sum, statchange_match_is_sum) %>% 
  mutate(className = factor(className,
                            levels = names(class_lookup),
                            labels = class_lookup),
         prop_improve = 1 - prop_decline,
         declineImprove = paste(n_decline,":",n_improve,sep = "")) %>% 
    # Remove NA for action in place (either threat is NA or missing info like AOH)
    filter(!(is.na(value)))

EDGE_sum <-
  alliucn %>% 
  filter(className %in% c("amphibia", "aves", "mammalia")) %>% 
  mutate(EDGE = case_when(is.na(EDGE.Rank) ~ FALSE,
                          !(is.na(EDGE.Rank)) ~ TRUE)) %>% 
  group_by(EDGE) %>% 
  summarise(n_neglect = length(which(actions_zero_sps)),
            n_nonNeglect = length(which(!(actions_zero_sps))),
            n = n(),
            prop_neglect = n_neglect / n,
            ED = median(ED, na.rm = TRUE))
```
_______________________________________________________________________________________

Rebecca A. Senior^1,2^*, Ruby Bagwyn^3^, Danyan Leng^4,5^, Alexander K. Killion^4,5^, Walter Jetz^4,5^, and David S. Wilcove^1,6^

^1^Princeton School of Public and International Affairs, Princeton University, Princeton, NJ, USA.

^2^Conservation Ecology Group, Department of Biosciences, Durham University, Durham DH1 3LE, UK

^3^Williams College, Williamstown, MA, USA.

^4^Department of Ecology & Evolutionary Biology, Yale University, New Haven, CT, USA.

^5^Center for Biodiversity and Global Change, Yale University, New Haven, CT, USA.

^6^Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA.

*__Corresponding author:__ rebecca.a.senior@gmail.com (R.A. Senior)

__ORCID iDs:__ orcid.org/0000-0002-8208-736X (R.A. Senior); orcid.org/0000-0003-4234-4154 (D. Leng); orcid.org/0000-0003-1449-8295 (A.K. Killion); https://orcid.org/0000-0002-1971-7277 (W. Jetz). 
  
__Authorship:__ R.A.S. and D.S.W. conceived the study. R.A.S., R.B., D.L., A.K.K., and W.J. collated and processed the data and W.J. contributed guidance on analysis. R.A.S. carried out the analyses and wrote the first draft of the manuscript, with substantial contribution from D.S.W. and R.B. All authors contributed to writing and editing of the manuscript. 

__Funding:__ High Meadows Foundation

__Key words:__ conservation effectiveness; IUCN Red List; extinction; conservation planning; biodiversity; threatened species

__Article type:__ Letter

__Number of words:__ abstract = 197; main text = 1933

__Number of References:__ 43

__Number of Figures:__ 4; __Number of Tables:__ 0 

# Abstract

Threatened species are, by definition, species in need of assistance. Absent the right conservation interventions, they are likely to disappear soon[@iucn_iucn_2020]. Globally, we have a poor understanding of how and where conservation interventions are applied, or how well they work[@sutherland_fresh_2018; @sutherland_building_2019]. Using information from the IUCN Red List and other global databases, we find that for species at risk from three of the biggest drivers of biodiversity loss -- habitat loss, overexploitation for international trade, and invasive species[@maxwell_biodiversity:_2016] -- many are seemingly lacking the appropriate types of conservation interventions. Indeed, while there has been substantial recent expansion of the protected area network, we still find that `r 100 - signif(matched_action_total$prop[matched_action_total$threat == "Habitat change"], 2)`% of threatened species have insufficient representation of their habitats within PAs. Conservation interventions are not implemented uniformly across different taxa and regions and, even when present, have infrequently led to substantial improvements in the species' status. For `r signif((length(which(alliucn$actions_zero_sps)) / nrow(alliucn[which(!(is.na(alliucn$actions_zero_sps))),])) * 100, 2)`% of the world's threatened terrestrial species, we find conservation interventions to be insufficient or absent. We cannot determine whether such species are truly neglected, or whether efforts to recover them are not included in major conservation databases. Without more and better targeted conservation efforts, the outlook for many of the world's threatened species is grim.

__197/200 words__

# Main text

The need for greater attention to biodiversity conservation is unequivocal and urgent[@diaz_pervasive_2019]. The world is entering its 6^th^ mass extinction event[@ceballos_accelerated_2015] - the first attributable to a single species: _Homo sapiens_. Biodiversity loss is a global concern and the focus of multiple international commitments, including those recently pledged in the Kunming-Montreal Global Biodiversity Framework (GBF) of the UN Convention on Biological Diversity. Nevertheless, species extinctions continue to accumulate[@wwf_living_2020]. 

Conservation efforts can forestall species extinctions[@bolam_how_2020; @monroe_dynamics_2019], but funding remains insufficient[@mccarthy_financial_2012; @waldron_targeting_2013]. Moreover, if effort is poorly targeted relative to risk[@christie_challenge_2020], fewer species will be saved than is otherwise possible. Prospects for biodiversity can be improved through more resources and more efficient allocation of what scarce resources are available. The latter requires that we identify which conservation interventions decrease species' risk of extinction[@sutherland_fresh_2018], along with what interventions have been implemented and where.

Until recently, little attention was given to assessing what works in conservation[@sutherland_fresh_2018; @burivalova_critical_2017]. Assessments of Protected Area (PA) effectiveness are a notable exception, with multiple studies finding that well-managed PAs mitigate biodiversity loss[@geldmann_effectiveness_2013; @geldmann_global-level_2019]. Similarly, extensive data demonstrate the benefit of invasive-species eradication efforts for island biota[@jones_invasive_2016]. There are, however, numerous other types of conservation interventions[@salafsky_standard_2008], many of which have yet to be assessed together and at scale.

The International Union for the Conservation of Nature (IUCN) conducts species-level assessments of extinction risk via its Red List of Threatened Species (hereafter: Red List)[@mace_quantification_2008; @iucn_iucn_2020]. Assessors also compile information on the conservation interventions implemented for each species. Studies have used these data to identify interventions associated with decreased extinction risk in birds and mammals[@luther_determinants_2016; @hayward_using_2011; @bolam_how_2020]. Here, we provide the first global assessment of patterns of conservation action across regions and taxonomic groups by supplementing the Red List data with a manual review of their 'Conservation Actions in Place' text, combined with four other databases on specific interventions: Map of Life[@jetz_include_2022]; the World Database on Protected Areas (WDPA)[@unep-wcmc_protected_2019]; Species+[@speciesplus_2021]; and the Database of Island Invasive Species Eradications (DIISE)[@spatz_globally_2017]. We consider all `r nrow(alliucn)` terrestrial threatened species (those classed as Vulnerable, Endangered or Critically Endangered) within taxonomic families that have been comprehensively assessed for the Red List (Table S1).

<!-- We overlay maps of conservation interventions (WDPA and DIISE) on threatened species' 'Area of Habitat' (AOH) maps, which we produce by refining range polygons according to species' habitat and elevation preferences[@jetz_projected_2007;@rondinini_habitat_2005].  -->

We initially focus on three of the greatest threats to biodiversity: habitat loss (including habitat degradation), overexploitation for international trade, and invasive species[@maxwell_biodiversity:_2016]. Each has a clearly matched conservation intervention: habitat protection (via PAs), trade control, and invasive-species control, respectively. While we recognise the potentially severe impact of domestic overexploitation[@liang_assessing_2023], we focus on international trade only, given greater availability of data on this threat and its matched intervention. 

Regarding habitat loss, we consider meaningful habitat protection to be in place only if a species-specific representation threshold of overlap is met, as defined by the Species Protection Score (SPS; see Methods). The SPS contributes to the Species Protection Index, a formal indicator of the UN GBF addressing species representation[@jetz_include_2022;@lefebvre_monitoring_2022]. Even where minimal representation within PAs is identified, we note that species may still be threatened by habitat loss where PAs are poorly managed or subject to downgrading, downsizing, or degazettement[@zeng_gaps_2023]. We also acknowledge the value of interventions besides PAs, including Other Effective Area-based Conservation Measures (OECMs)[@gurney_biodiversity_2021], spatial planning[@green_farming_2005], and habitat restoration. However, we lack comparably comprehensive data to include them here. 

<!-- None of the interventions we investigate is necessarily sufficient in isolation, although each is an appropriate, widely-accepted minimum response to the threat in question.  -->

For species threatened by any of the three major drivers of biodiversity loss outlined above, we begin with two key questions. (1) What proportion of species receive the appropriate type(s) of conservation intervention? (2) Does a species' taxonomy, biogeography or extinction risk influence the likelihood that the appropriate intervention will be made? Finally, we consider all categories of in situ conservation interventions documented by the Red List[@salafsky_standard_2008] - additionally including reintroduction, international legislation and education - to identify which threatened species have no documented conservation interventions. For species lacking documented conservation interventions to tackle their main threats, and for those lacking any documented conservation interventions, we ask: is the apparent lack of conservation attention qualitatively associated with changes in species' status on the Red List?

<!-- Few small-ranged species have sufficient protection of their remaining habitat [@pimm_biodiversity_2014; @pimm_how_2018] -->

Globally, we find significant shortcomings in documented conservation interventions for the world's threatened terrestrial species. Most threatened species at risk of overexploitation for international trade are documented as being subject to international trade control (`r signif(matched_action_total$prop[matched_action_total$threat == "Overexploitation"], 2)`%). However, for those under threat from habitat loss, only `r signif(matched_action_total$prop[matched_action_total$threat == "Habitat change"], 2)`% have sufficient representation of their habitat in PAs to meet target thresholds. As noted elsewhere, small-ranged species are particularly poorly represented in the current PA network[@pimm_biodiversity_2014; @pimm_how_2018; @zeng_gaps_2023]. Among species threatened primarily by invasive species, only `r signif(matched_action_total$prop[matched_action_total$threat == "Invasive species"], 2)`% are documented as receiving invasive species control (Figure \@ref(fig:fig-1)). If we relax the criterion for habitat protection to include mere occurrence in at least one PA, we find that `r signif(matched_action_hc_targetRL$prop,2)`% of species threatened by habitat loss are covered (Figure S2-S3). Some taxonomic biases also exist in the extent to which species' threats are documented as being appropriately addressed. For example, `r signif(matched_action_class$prop[matched_action_class$className == "Magnoliopsida"][1], 2)`% and `r signif(matched_action_class$prop[matched_action_class$className == "Magnoliopsida"][3], 2)`% of flowering plants in the class Magnoliopsida (Figure \@ref(fig:fig-1)) are documented as receiving meaningful habitat protection or invasive-species control, respectively (Figure \@ref(fig:fig-1)), whereas `r signif(matched_action_class$prop[matched_action_class$className == "Aves"][1], 2)`%, `r signif(matched_action_class$prop[matched_action_class$className == "Aves"][2], 2)`% and `r signif(matched_action_class$prop[matched_action_class$className == "Aves"][3], 2)`% of birds are documented as receiving the appropriate interventions to tackle habitat loss, international trade, or invasive species, respectively.

```{r fig-1, fig.cap = "The proportion of threatened species documented as  receiving the appropriate type of conservation intervention to tackle three major threats to biodiversity. The threats are: habitat loss (a), overexploitation for international trade (b), and invasive species (c). Different bar colours denote different Red List categories: Vulnerable (VU; yellow), Endangered (EN; orange), and Critically Endangered (CR; red). Small numbers above the bars represent the total number of species included in our analyses, by taxonomic class and threat. __75/100 words__", fig.width = 10/2.54, fig.height = 16/2.54}
levels(matched_action_sum$threat) <- 
  c(expression(paste(bold('a')," Habitat loss")), 
    expression(paste(bold('b')," International trade")),
    expression(paste(bold('c')," Invasive species")))
classes <- levels(matched_action_sum$className)
labels <- paste(
  "<img src=", 
  "'",
  file.path('figs/phylopic', paste(tolower(classes), 'png', sep = ".")),
  "' ",
  "width='10' />", 
  classes,
  sep = ""
)
names(labels) <- classes

p <-
  ggplot(data = matched_action_sum) +
  facet_wrap(~ threat, ncol = 1,
             labeller = label_parsed) +
  geom_text(aes(x = className,
                y = 1.06,
                label = classN),
            size = 2.2,
            colour = "transparent") +
  geom_rect(aes(xmin = classNumeric - 0.5,
                xmax = classNumeric + 0.5,
                ymin = -Inf,
                ymax = Inf,
                fill = className)) +
  scale_fill_manual(values = rep(c("grey80", "grey92"),4), 
                    guide = "none") +
  geom_text(aes(x = className,
                y = 1.06,
                label = classN),
            size = 2.2,
            colour = "black") +
  new_scale_fill() +
  geom_bar(aes(x = className, 
               y = nT / N,
               fill = factor(iucn)),
           stat = "identity",
           position = position_dodge(width = 0.9))  +
  scale_fill_manual(labels = c("VU", 
                               "EN", 
                               "CR"),
                    name = NULL,
                    values = iucnPal) +
  ylab("Proportion of species with\nappropriate intervention (%)") +
  myTheme() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_markdown(color = "black",
                                       angle = 45, 
                                       hjust = 1, vjust = 1,
                                       margin = unit(c(-5, 0, 0, 0), "pt")),
        plot.margin = margin(0.1,0.1,0,0.1, unit = "cm"),
        strip.text = element_text(size = 10, hjust = 0),
        legend.key.size = unit(0.4, "cm"),
        legend.margin = margin(0,0,-0.3,0,"cm"), 
        legend.position = "top") +
  scale_y_continuous(limits = c(0, 1.3),
                     breaks = seq(0, 1, 0.2),
                     labels = seq(0, 100, 20),
                     expand = c(0,0)) +
  scale_x_discrete(labels = labels)

sub_imgs <- imgs[match(tolower(levels(matched_action_sum$className)), names(imgs))]

for (i in 1:length(sub_imgs)) {
  x <- (1:length(sub_imgs))[i]
  p <- p +
    add_phylopic(img = sub_imgs[[i]], 
                 alpha = 1, x = x, y = 1.2, ysize = 0.15)
}
p
```

The distribution of species lacking appropriate types of conservation intervention shows considerable geographic variability (P < 0.001; Figure \@ref(fig:fig-2); Figure S8-S11). Several regions contain high numbers of amphibians requiring invasive-species control (P < 0.001), including Madagascar, Central America and Australia (Figure \@ref(fig:fig-2)c). The majority of cases (`r signif((sort(invspp_neglect_sum_amphib, decreasing = TRUE) / sum(invspp_neglect_sum_amphib) * 100)[1], 2)`%) where control of invasive species is needed but lacking pertain to a lack of control of chytrid fungus, _Batrachochytrium dendrobatidis_, for which there is not yet an effective treatment[@geiger_antifungal_2017].

With respect to habitat loss, amphibians lack meaningful habitat protection in Central America (Figure \@ref(fig:fig-2)a), while mammals are notably lacking in habitat protection in Indonesia (Figure \@ref(fig:fig-2)g) as are birds in both South and Central America and Indonesia (Figure \@ref(fig:fig-2)d). Most species threatened by international trade are documented as receiving some international trade control (Figure \@ref(fig:fig-1)b), though relatively high numbers of exploited birds in Indonesia seemingly lack such protection (Figure \@ref(fig:fig-2)e; Figure S8). 

(ref:cap-2) The total number of threatened species within each country apparently lacking the appropriate type of conservation intervention for three major threats. Results are summarised for each of three major threats to biodiversity with a clearly matched intervention (columns), and for each taxonomic group (rows). For clarity, we include here only the three vertebrate classes that have range data and have been comprehensively assessed, but see Figure S8 for full results across all assessed families. __74/100 words__

```{r fig-2, fig.cap = "(ref:cap-2)", fig.width = 18.3/2.54, fig.height = 9/2.54}
fig2 <- readPNG(file.path("figs/maps", "match_ctry_total_map_sub.png")) %>% 
  rasterGrob(interpolate = TRUE)
ggplot() +
  theme_void() +
  # theme(plot.margin = margin(-1,-1,-1,-1, unit = "cm")) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  annotation_custom(fig2,
                    xmin = -Inf, xmax = Inf,
                    ymin = -Inf, ymax = Inf)
```

Finally, we assessed how many of the threatened species in our database are not documented as receiving meaningful habitat protection or any of the other categories of conservation intervention, expanded to include additional measures such as reintroduction and education[@salafsky_standard_2008]. Overall, we find that `r signif((length(which(alliucn$actions_zero_sps)) / nrow(alliucn[which(!(is.na(alliucn$actions_zero_sps))),])) * 100, 2)`% (`r length(which(alliucn$actions_zero_sps))`/`r nrow(alliucn[which(!(is.na(alliucn$actions_zero_sps))),])`) of threatened terrestrial species lack meaningful habitat protection or any other documented conservation interventions (Figure \@ref(fig:fig-3)). We emphasize that this percentage is derived in part based on the Species Protection Score, which uses a species-specific representation threshold to determine if a species has meaningful representation in PAs[@jetz_include_2022]. When we relax that threshold to whether or not a species occurs in any PA to any extent (as noted by the Red List), the proportion falls to `r signif((length(which(alliucn$actions_zero_targetRL)) / nrow(alliucn[which(!(is.na(alliucn$actions_zero_targetRL))),])) * 100, 2)` %
(`r length(which(alliucn$actions_zero_targetRL))`/ `r nrow(alliucn[which(!(is.na(alliucn$actions_zero_targetRL))),])`; Figure S4-5). Taxa such as amphibians fare worse than others (Deviance = `r round(results_actionZero_class$actions_zero_sps$Deviance[2], 0)`, DF = `r round(results_actionZero_class$actions_zero_sps$Df[2], 0)`, `r siglevel(results_actionZero_class$actions_zero_sps[,5][2])`), with `r signif(filter(zero_action_class, className == "Amphibia")$prop,2)`% lacking meaningful habitat protection or any other documented conservation interventions, compared to `r signif(filter(zero_action_class, className == "Aves")$prop, 2)`% for threatened birds. Taxonomic biases likely result from increased attention to charismatic and easily-studied species[@colleony_human_2017; @christie_challenge_2020;@troudet_taxonomic_2017], which also translates to more frequent Red List assessments. Across some (but not all) taxa, species at greatest risk of extinction are more likely to have documented attention (Deviance = `r round(results_actionZero_iucn$actions_zero_sps$Deviance[2], 0)`, DF = `r round(results_actionZero_iucn$actions_zero_sps$Df[2], 0)`, `r siglevel(results_actionZero_iucn$actions_zero_sps[,5][2])`). This corroborates Luther et al.[@luther_determinants_2016], who concluded that species at greater risk of extinction receive more conservation attention.

```{r fig-3, fig.cap = "The proportion of threatened species lacking meaningful habitat protection, or lacking any of the other six categories of documented conservation interventions (irrespective of threat). Different bar colours denote different Red List categories: Vulnerable (yellow), Endangered (orange), and Critically Endangered (red). Small numbers above the bars represent the total number of species included in our analyses, by taxonomic class. __58/100 words__", fig.width = 8.9/2.54, fig.height = 8.9/2.54}
classes <- levels(zero_action_sum$className)
labels <- paste(
  "<img src=", 
  "'",
  file.path('figs/phylopic', paste(tolower(classes), 'png', sep = ".")),
  "' ",
  "width='10' />", 
  classes,
  sep = ""
)
names(labels) <- classes

p <-
  ggplot(data = zero_action_sum) +
  geom_text(aes(x = className,
                y = 1.08,
                label = classN),
            size = 2.2,
            colour = "transparent") +
  geom_rect(aes(xmin = classNumeric - 0.5,
                xmax = classNumeric + 0.5,
                ymin = -Inf,
                ymax = Inf,
                fill = className)) +
  scale_fill_manual(values = rep(c("grey80", "grey92"),4), 
                    guide = "none") +
  geom_text(aes(x = className,
                y = 1.03,
                label = classN),
            size = 2.2,
            colour = "black") +
  new_scale_fill() +
  geom_bar(aes(x = className, 
               y = nT / N,
               fill = factor(iucn)),
           stat = "identity",
           position = position_dodge(width = 0.9))  +
  scale_fill_manual(labels = c("VU", 
                               "EN", 
                               "CR"),
                    name = NULL,
                    values = iucnPal) +
  ylab("Proportion of species with\nno interventions (%)") +
  myTheme() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_markdown(color = "black",
                                       angle = 45, 
                                       hjust = 1, vjust = 1,
                                       margin = unit(c(-5, 0, 0, 0), "pt")),
        legend.key.size = unit(0.4, "cm"),
        legend.margin = margin(0,0,-0.3,0,"cm"), 
        legend.position = "top") +
  scale_x_discrete(labels = labels) +
  scale_y_continuous(limits = c(0, 1.17),
                     breaks = seq(0, 1, 0.2),
                     labels = seq(0, 100, 20),
                     expand = c(0,0))


sub_imgs <- imgs[match(tolower(levels(droplevels(zero_action_sum$className))), names(imgs))]

for (i in 1:length(sub_imgs)) {
  x <- (1:length(sub_imgs))[i]
  p <- p +
    add_phylopic(img = sub_imgs[[i]], 
                 alpha = 1, x = x, y = 1.1, ysize = 0.09)
}
p
```

While many threatened species are beneficiaries of documented conservation attention, a critical question is: do those interventions work? From 2006-2020, `r table(statchange$cat_change_bi)[1]` species were uplisted to a higher threat category, and `r table(statchange$cat_change_bi)[2]` species were downlisted to a lower threat category due to a genuine decrease in extinction risk (Figure S15). Of the downlisted species, only `r signif(100*(length(which(statchange$cat_change_bi == 1 & statchange$actions_zero_sps)) / length(which(statchange$cat_change_bi == 1))),2)`% (`r length(which(statchange$cat_change_bi == 1 & statchange$actions_zero_sps))`) lacked any documented conservation interventions. Alarmingly, `r signif(100*(length(which(statchange$cat_change_bi == 0 & !(statchange$actions_zero_sps))) / length(which(statchange$cat_change_bi == 0))),2)`% (`r length(which(statchange$cat_change_bi == 0 & !(statchange$actions_zero_sps)))`/`r length(which(statchange$cat_change_bi == 0))`) of uplisted species had received at least some documented conservation attention, suggesting that the measures employed were insufficient to reverse declines. Focusing on species facing only one major threat (habitat loss, international trade, or invasive species), a higher proportion were uplisted when the appropriate intervention was apparently lacking, although this was not consistently the case (Figure \@ref(fig:fig-4)). Thus, while conservation interventions are qualitatively associated with improvements in species' Red List status, corroborating that conservation can succeed[@bolam_how_2020; @monroe_dynamics_2019], the mere existence of 'something rather than nothing' is not sufficient to reverse declines. Previous studies have documented large variation in how well conservation interventions are implemented, such as variation in PA management[@leverington_global_2010]. Additionally, ecological time lags occur in response to both positive and negative change[@watts_ecological_2020], and there are time lags inherent to the Red List process itself. Species must have met the criteria for a lower threat category for at least 5 years before the status change is implemented[@iucn_criteria_2012]. Few taxa were reassessed in the period 2006-2020, and birds were reassessed more frequently (approximately every four years) than any other group.

Given the geographic patterns in conservation attention, it is possible that certain groups of species are disproportionately neglected. We found a weak trend whereby 'Evolutionary Distinctiveness' - a measure of species' relative contribution to phylogenetic diversity[@redding_incorporating_2006] - was lower in species lacking meaningful habitat protection or any other documented conservation interventions, compared to species with at least one documented intervention (Figure S16; Deviance = `r round(results_actionZero_ED$Deviance[2], 0)`, DF = `r results_actionZero_ED$Df[2]`,  P = 0.05). The total number of endemic threatened species with documented conservation interventions was positively correlated with the number that lacked such interventions (Figure S13). This may suggest that apparent neglect does not result from lack of will, but rather from insufficient capacity to act or to report for countries with greater numbers of threatened endemic species. However, the relationship was not straightforward (Figure S12), as the probability of apparent neglect for any given threatened endemic species was not statistically associated with country GDP (Gross Domestic Product; Deviance = `r round(probZero_country_gdp$Deviance[2], 0)`, DF = `r probZero_country_gdp$Df[2]`, `r siglevel(probZero_country_gdp[,5][2])`) or total number of endemic threatened species (Deviance = `r round(probZero_country_n$Deviance[2], 0)`, DF = `r probZero_country_n$Df[2]`, `r siglevel(probZero_country_n[,5][2])`; Figure S14).

Teasing apart observed geographic patterns (Figure \@ref(fig:fig-3)) to discern causality requires a more nuanced assessment of conservation resources, one that additionally accounts for international aid and the sufficiency of interventions[@leverington_global_2010]. It is also important to note that even with detailed assessment guidelines and extensive training, geographic and taxonomic biases in assessor reporting exist[@hayward_ambiguity_2015]. This does not undermine our results, because it is crucial that scientists can identify where improved documentation is needed. It does, however, complicate interpretation, because we are not currently able to distinguish between the need for documentation versus true absence of conservation interventions.

```{r fig-4, fig.cap = "The percentage of species downlisted (blue) versus uplisted (red) to another threat category. We exclude 'non-genuine' status changes, which result from revisions in taxonomy or corrections of erroneous assessments. Columns show results for species with any one of habitat loss (a), overexploitation for international trade (b) or invasive species (c) listed as a major threat, where the appropriate type of conservation intervention is either documented as in place (left) or not (right). Bold numbers represent the number of species changing Red List status, comprising the number downlisted (down arrow) plus the number uplisted (up arrow). __95/100 words__", fig.width = 16/2.54, fig.height = 10/2.54}

# Summarise status change data by class & conservation action
statchange_match_sum2 <-
    gather(statchange_match_sum, key = "direction", value = "n_change",
           "n_improve", "n_decline") %>%
    arrange(className, value, matched_action, direction) %>%
    mutate(prop = n_change / n)
# Write to CSV to use when creating figure manually in Powerpoint
write.csv(arrange(statchange_match_sum, className, matched_action, value),
          "figs/Fig4.csv", row.names = FALSE)

fig4 <- readPNG("figs/Fig4.png") %>% 
  rasterGrob(interpolate = TRUE)
ggplot() +
  theme_void() +
  # theme(plot.margin = margin(-1,-1,-1,-1, unit = "cm")) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  annotation_custom(fig4,
                    xmin = -Inf, xmax = Inf,
                    ymin = -Inf, ymax = Inf)
    
```

Going forward, IUCN's Green Status of Species, which aims to track species recovery[@akcakaya_quantifying_2018], coupled with the expanding Conservation Evidence database[@sutherland_building_2019], should enable researchers to better explore the positive trends of some species on the Red List. Moreover, the commitment by over 50 nations to expand the global PA network to encompass at least 30% of terrestrial and marine ecosystems by 2030 provides an opportunity to protect those species threatened by habitat loss that are currently under-represented in PAs[@30by30_2021]. Complementary-based approaches that efficiently ensure adequate species representation, such as those delivered by the Half-Earth Project, will be key to complement existing approaches for area-based conservation [@rinnan_multi-scale_2021]. OECMs will play a significant role in conserving threatened species outside of traditional PAs[@gurney_biodiversity_2021; @sze_reduced_2022], especially on lands owned and managed by indigenous communities. However, a critical first step in improving global conservation practices is documenting what we are already doing. Globally, there are taxonomic and geographic biases in the species documented as receiving conservation attention, with limited instances of species being downlisted to lower categories of threat. Conservation can succeed, but without more and better-targeted investment, we risk surrendering the world's threatened species to mass extinction.

__1933/1500 words__

# Methods

## Data sources

Information on species' current status, threats and conservation interventions was downloaded from the Red List on 14^th^ July 2020[@iucn_iucn_2020]. We focused only on terrestrial or terrestrial and freshwater species (as defined on the Red List) that are classified as threatened (Vulnerable, Endangered or Critically Endangered) and fall within the taxonomic groups that have been comprehensively assessed to at least family level (Table S1): amphibians (2220), birds (1295), cycads (192), mammals (1258), conifers (198), crocodiles and alligators (9), birches (19), magnolias (148), southern beeches (11), teas (88), cacti (415), freshwater crabs (36) and chameleons (74). This gave a total of `r nrow(alliucn)` species. We reclassified threats into six broader categories (Table S2), in line with those listed by the Convention on Biological Diversity (CBD)[@hirsch_global_2010]:

1. Habitat loss (including habitat degradation)
2. Overexploitation
3. Pollution
4. Invasive alien species
5. Climate change
6. Other (e.g., unknown threats and threats due to natural causes, such as volcanic eruptions)

<!-- * Trade check different to Eyal Frank because no manual search, just relying on IUCN Red Listed species that are intentionally & internationally traded -->
<!--   * With our criteria, 560 threatened species threatened by international trade; relaxing to include all scopes/severities/timings increases to 729 which is still less than Eyal Frank's 958 -->

For the broad category of habitat loss, we further required that the threat be coded, according to the Red List 'stresses' classification scheme, as an "Ecosystem/Community Stress". This includes ecosystem stresses in the form of ecosystem conversion, ecosystem degradation, and indirect ecosystem effects (such as ecosystem fragmentation), but does not include stresses that solely and directly affect the species. We applied this additional restriction because some of the threats that could potentially drive habitat loss, such as "Recreational activities" (Table S2), could instead cause species stresses such as direct mortality, competition and reduced reproductive success, while leaving habitat intact.

We restricted our analysis of the CBD threat 'overexploitation' to cases involving the intentional use of internationally traded species because this is a major known threat to biodiversity that can be clearly matched to the conservation intervention 'international trade control'[@frank_long_2019; @morton_impacts_2021]. Other forms of resource use, such as "local, subsistence hunting", fall into the threat category 'Other'. 

In all analyses, we focus only on threats that we deemed to have a major influence on a species' risk of extinction, based on the IUCN-defined fields: 'scope', 'severity' and 'timing'. Specifically, the threat must : (1) affect more than 50% of the global population of a species; (2) have caused "Slow, Significant Declines", "Rapid Declines", or "Very Rapid Declines" in the population; and (3) have occurred in the past or be ongoing (i.e., not a predicted future threat). We also include cases where the scope or severity of the threat is unknown because, depending on the threat category, between `r min(threat_unknown_count)` and `r max(threat_unknown_count)` threatened species had one of the six threats listed but with unknown scope and/or severity. Excluding cases where threat severity or scope was unknown did not affect our conclusions (Figure S6-S7).

Of the `r length(which(alliucn$iucn == 5))` species in our database that are categorised on the Red List as Critically Endangered, `r length(which(alliucn$iucn == 5 & alliucn$possiblyExtinct))` are possibly extinct. Species that are possibly extinct are potentially less likely to have conservation interventions in place, since conservationists may not know where the species occurs or what interventions it needs. Moreover, such species may be subject to triage due to the low chance of conservation success. However, we found that our results were robust to excluding possibly extinct species (Figure S6-S7).

For each species, the Red List denotes various conservation interventions as either in place or not. Our focus was on the following six interventions, as they are defined in the Red List 'conservation actions in place' classification scheme:

1. Does the taxon occur in at least one Protected Area?

2. Is the taxon subject to any international management/trade controls?

3. Is there invasive species control or prevention?

4. Has the taxon been successfully reintroduced or introduced benignly?

5. Is the taxon included in international legislation?

6. Is the taxon the subject of any recent education or awareness programmes?

We did not consider the intervention 'ex situ conservation', in place for `r table(alliucn$exsitu)[2]` species in our dataset, because we are interested in actions focused on decreasing species' risk of extinction in the wild (although we recognise that ex situ conservation can ultimately contribute to the recovery of threatened species in the wild).

Although we have adopted the Red List classification scheme here, and we use Red List data to detect the presence of most conservation interventions, in our main analyses we chose not to use Red List data to determine whether or not a species occurs in PAs. The Red List's binary classification of whether a species is present in PAs is determined by published and unpublished literature, in combination with expert knowledge of the distribution of populations, rather than by employing a representation threshold. This allows a species to be categorised as receiving habitat protection if just a small fraction of its habitat falls within the boundaries of a single PA. Instead, we use the Species Protection Score, which is used in the calculation of Species Protection Index - the formal UN Biodiversity indicator of species representation[@jetz_include_2022;@lefebvre_monitoring_2022] (Figure S1). Species Protection Scores are calculated as the percentage of each species' range that occurs within the boundaries of Protected Areas (World Database on Protected Areas; WDPA)[@unep-wcmc_protected_2019], relative to a pre-determined, species-specific representation threshold. For example, if 50% of a species range occurs within PAs and the representation threshold for that species is 50%, then the Species Protection Score is 100. Conversely, if the representation threshold for that species is 80%, then a 50% overlap of the species' range with PAs would correspond to a Species Protection Score of only 62.5 (i.e. (50/80) x 100). While work on a more ecologically fine-tuned yet broadly applicable determination of representation thresholds is in progress, the threshold itself is informed  by Rodrigues et al.[@rodrigues_global_2004], whereby species with less than or equal to 10,000 km^2^ habitat must have 100% of that habitat occurring within PAs for the species to be considered meaningfully represented in PAs. Species with more than or equal to 250,000 km^2^ habitat must have at least 15% of that habitat occurring within PAs for the species to be considered meaningfully protected by PAs. For all other values of range size (10,000 km^2^ &le; range &le; 250,000 km^2^), a log-linear interpolation between 15 and 100 is applied to calculate the appropriate representation threshold. These thresholds assume that species with less habitat require a greater proportion of that habitat to occur within PAs for those species to be considered protected from habitat loss and degradation. 

In this study, we assigned a binary value to the Species Protection Score such that the species was considered to be meaningfully represented in PAs only if its Species Protection Score was 100. All other values of the Species Protection Score indicate a failure to achieve meaningful representation of that species within the existing PA network. We note that the calculation of Species Protection Scores can be performed in two ways, depending on the range maps available. All species on Map of Life have Expert Map Ranges, and hence all species had Species Protection Scores derived from overlapping these Expert Map Ranges with PAs. In addition, most species also had Species Protection Scores based on overlapping PAs with 'Habitat-Suitable Ranges' (HSR), whereby each species' range is refined according to its habitat and elevation preferences [@powers_global_2019] (also known as 'Area of Habitat'). To be conservative, we considered a species to be meaningfully represented in PAs if either of these Species Protection Scores was 100. To explore species HSR results and Protection Scores, see mol.org/indicators/protection. 

For independent validation of the Species Protection Score, we also applied our own analogous protocol to derive HSR for each species and calculate its percentage overlap with the WDPA. Species-specific representation thresholds were calculated as above. HSR could be calculated for most of the species in our dataset, because we focus on the taxonomic families that have spatial range data available from the Red List and BirdLife International[@birdlife_2020; @iucn_iucn_2020], including all bird, mammal and amphibian families, plus select families of reptiles (Chamaeleonidae, Crocodylidae, Gavialidae), and flowering plants (Theaceae, Magnoliaceae, Betulaceae, Nothofagaceae). The WDPA was cleaned following best-practice guidelines[@butchart_shortfalls_2015; @protectedplanet_best-practice_2020]. All spatial analyses were conducted in Google Earth Engine[@gorelick_google_2017]. Following, Powers & Jetz (2019)[@powers_global_2019] and Brooks et al. 2019[@brooks_measuring_2019], HSR was determined by refining range polygons to areas with suitable land cover and elevation. Habitat preferences and elevation limits for each species were obtained from the Red List[@iucn_iucn_2020] and Quintero and Jetz 2018[@quintero_global_2018]. For elevation we use the EarthEnv Digital Elevation Model Version 1[@robinson_earthenv-dem90:_2014], resampled from ~90 m to 1 km. For land cover we use a map of terrestrial habitat types for the year 2015, specifically designed to match the habitat classification scheme of the IUCN[@jung_global_2020]. Where a species' habitat preference is given only to the coarser level 1 classification (e.g. 'Forest'), we conservatively assume that all nested level 2 categories are also suitable. We use the fractional habitat types map aggregated to 1 km resolution, following the recommendation of Jung et al.[@jung_global_2020].

In our calculation of HSR, limitations of the input data resulted in 621 species with zero HSR. In these cases, we additionally verified if that species occurs in PAs if it had been observed within a PA in the last five years (Figure S1), based on point occurrence records from the Global Biodiversity Information Facility (GBIF)[@gbif_2021]. GBIF records were retrieved in R version 4.1.0[@r_core_team_2021], using the packages `rgbif`[@chamberlain_r_2017] and `taxize`[@chamberlain_taxize_2013], and points were buffered by 300 m to allow for positional errors[@jung_global_2020]. By repeating the analyses of the main text with our independent calculation of HSR, we find very similar results to the main analyses based on the Species Protection Score (Supplementary Figures 2-5). However, if we use a less conservative approach whereby the determination of whether or not a species occurs in PAs is based on both the Red List data and threshold percentage overlap of HSR with PAs (Supplementary Figures 2-5), a far greater number of species appear to have meaningful habitat protection, and therefore a greater number also appear to have at least one documented conservation intervention.

Preliminary checks suggested that conservation intervention information is lacking for many assessed species in the Red List data download, despite being evident in the detailed text description of species' Red List profiles. As a result, we supplemented the binary classification provided in the tabular data download by manually reviewing the 'Conservation Actions In Place' text for each threatened species (n = `r nrow(alliucn)`), using the same criteria as defined by IUCN[@iucn_conservation_2012]. Thus, the first test of whether a conservation intervention is in place is whether it is recorded as such in the Red List tabular data. If not, the second test is whether the intervention is described as in place in the Red List text description. Failing both tests, the conservation intervention is recorded as not in place, except for the subset of interventions with additional or alternative sources of information outside of the Red List. These interventions were: PA coverage (as described above), international trade control and legislation, and invasive species control. The details of the additional tests are described in detail below and in Figure S1. 

International legislation includes international trade control, hence only one was used in any given analysis (i.e., either international legislation or international trade control). Most of our analyses focus on interventions matched to one of three major threats, hence we use international trade control as the appropriate intervention for the threat of overexploitation for international trade. We consider a species to be subject to international trade control only if stated as such on the Red List, or if the species (or a larger taxonomic group of which it is part) is listed on any of the appendices of CITES (Convention on International Trade in Endangered Species of Wild Fauna and Flora; Figure S1)[@speciesplus_2021]. 

Several additional conservation interventions exist that are less targeted to a specific threat, but which are, nevertheless, important tools for reducing species' overall risk of extinction[@salafsky_standard_2008]. Thus, we include an additional analysis of how many species have *any* documented conservation interventions, irrespective of what their main threats are. The scope of this analysis is broader, hence we replace the intervention of international trade control with the broader intervention of international legislation, which additionally includes multilateral agreements that are not directly concerned with trade control, such as CMS (the Convention on the Conservation of Migratory Species of Wild Animals). Specifically, we consider species to be subject to international legislation if they meet any one of the following criteria (Figure S1): 

1. The species is subject to international trade control as defined above (stated as such on the Red List, or the species is listed on any of the CITES appendices). 
2. The species is listed on published legislation listings from Species+[@speciesplus_2021], which includes all species that are listed in the Appendices of CITES and CMS, as well as species listed in the Annexes to the EU Wildlife Trade Regulations. 
3. The species is described as subject to a named multilateral agreement (see Table S4 for legislation considered). 

In addition to the described calculation of HSR overlap with PAs, we also overlaid HSR with spatial data available for invasive species eradication efforts (Database of Island Invasive Species Eradications: DIISE)[@spatz_globally_2017]. For island species at risk from invasive vertebrates, we identified cases where any of the threatened species' HSR overlapped with an island from which its threatening invasive species had been successfully eradicated (Figure S1)[@spatz_globally_2017]. This determination was based only on eradication of vertebrate species specifically named by the Red List as impacting the threatened species in question. Threatened island species were identified as those threatened species with more than 95% of their range area occurring on islands, based on the overlap between island polygons from the Global Island Database[@unep-wcmc_global_2015], and species' historical range (the sum of polygons with presence codes: 'extant', 'probably extant', 'possibly extinct', extinct' and 'presence uncertain')[@birdlife_2020]. We focus on threatened island species only, because spatial data for invasive species eradication efforts are reliably available only for islands[@spatz_globally_2017]. We note that wide-ranging threatened species may also have part of their range on islands, where they, too, may be threatened by invasive island vertebrates. However, with our data sources it is not possible to determine precisely where in a species' range a named invasive vertebrate is exerting its impact, and thus we can only be sure that a threatened species benefits from eradication efforts on islands if that threatened species is itself an island endemic.

To determine if there are geographic hotspots of apparent conservation neglect, we mapped the spatial distribution of threatened species that lack documented conservation interventions. We determined the countries in which each species occurs using information from Map of Life. The steps described above identified the species lacking appropriate types of conservation intervention to tackle the three major threats that we assessed, as well as the species lacking any documented conservation interventions. In both cases (separately), we created maps by summing the number of apparently neglected species occurring in each country. In statistical analyses we focus on country endemics to allow us to more precisely pinpoint where conservation effort is apparently lacking. We include all species in the maps (except for Figure S12), but note that the conservation interventions data provided by the Red List are not spatially explicit. As such, the presence or absence of conservation interventions is assigned to all parts of a species' global distribution. For non-endemics, there are instances where a species appears to lack a particular intervention in all countries where it occurs, but in fact the intervention is only necessary in a subset of countries. Conversely, there are instances where a species appears to benefit from an intervention throughout its entire range, but in reality the intervention is implemented in only one of the countries in which it occurs. 

To qualitatively explore whether documented conservation interventions were associated with species' risk of extinction, we focused on species that have changed Red List status. Any official change in Red List status is the result of extensive assessment by IUCN. Status change tables covering the years 2006-2020 were downloaded on 14^th^ July 2020[@iucn_statchange_2020]. We excluded records where the change in status of a given species was due to non-biological factors, such as new information or a change in taxonomy. We also excluded cases where an allegedly 'genuine' status change was later superseded by a non-genuine change of status in the opposing direction. For example, in 2008 the Mauritian flying fox, _Pteropus niger_, was listed as having genuinely improved in status, moving from Endangered to Vulnerable, but in 2013 the species was uplisted back to Endangered because previous assessments were found to have used incorrect data. For species genuinely changing status multiple times, we use only the first instance.

<!-- for birds, @wilman_eltontraits_2014 for mammals and @oliveira_amphibio_2017 for amphibians. Reptile body mass was calculated from snout-vent length using data and allometric equations from @meiri_traits_2018. We identified island birds using data from @tobias_integrating_2019, whereby a species was classed as island dwelling if >25% range occurred on islands. We followed this same protocol to identify island mammals, amphibians and reptiles,  -->

<!-- * Phylogenies from @jetz_global_2012 for birds, @kuhn_simple_2011 for mammals, @jetz_interplay_2018 for amphibians and @tonini_fully-sampled_2016 for reptiles -->

## Analyses

All analyses were performed in R version 4.3.2[@r_core_team_2021]. Full model results are reported in Supplementary Tables 6-9. Model inference was made using Likelihood Ratio Tests, dropping each variable in turn and comparing the reduced model to the full model[@zuur_mixed_2009]. We used generalized linear models (GLMs) with a binomial error structure, fit using the `glm` function of the `lme4` package[@bates_fitting_2015], to model both the proportion of species documented as receiving the appropriate type of conservation intervention, and the proportion of species with no documented interventions. In the former, each threat with a matched conservation intervention was modelled separately, to avoid double-counting species facing multiple threats. Explanatory variables were taxonomic class and Red List category, and the interaction between them. 

We additionally tested whether the number of species documented as receiving conservation attention differed between countries. The two response variables tested were: the proportion of species in each country documented as receiving the appropriate type of conservation intervention (tested for each threat separately); and the proportion of species in each country with no documented interventions. Both response variables were modelled against the explanatory variable of country, again using GLMs with a binomial error structure. Subsequently, to explore the drivers of country-level conservation effort, we tested whether the modelled probability of a species receiving no documented conservation interventions was predicted by country Gross Domestic Product (GDP[@worldbank_2021]) or the total number of threatened endemic species occurring in that country.

We might expect that instances of species changing Red List status would be associated with both: (1) the presence/absence of the type of conservation intervention appropriate to tackle the three major threats that we assessed; and (2) the presence/absence of any documented conservation interventions. In the first case, we considered the subset of species changing Red List status that had either habitat loss, overexploitation for international trade, or invasive species, as their only major threat of the three threats that we assessed. We summarised the number of species uplisted and downlisted according to whether or not they were documented as receiving the appropriate type of conservation intervention to tackle the one major threat they faced (out of the three threats that we assessed). We consider the appropriate type of intervention for these three major threats to be (respectively): occurring within a PA; international trade control; and invasive-species control. For the second question, we summarised the number of species being uplisted to a higher threat category, or downlisted to a lower threat category, according to whether any conservation interventions had been documented for those species.

Finally, we explored the potential consequences on phylogenetic diversity of apparent biases in conservation effort, by modelling the proportion of all amphibian, bird and mammal species with and without documented conservation interventions, according to their Evolutionary Distinctiveness (ED)[@redding_incorporating_2006]. ED is a measure of species' relative contribution to phylogenetic diversity[@redding_incorporating_2006]. ED data were downloaded from the EDGE of Existence website[@edge_2021]. We used a GLM with a binomial error structure, including both Red List category and ED as explanatory variables. The latter was standardised within each taxonomic class to range between 0 and 1.

## R Packages 

Data wrangling: `dplyr`[@dplyr_2020], `pdftools`[@pdftools_2020], `purrr`[@purrr_2020], `tidyr`[@tidyr_2020]. GBIF data download: `rgbif`[@chamberlain_r_2017], `taxize`[@chamberlain_taxize_2013]. Spatial data: `raster`[@raster_2020], `sf`[@sf_2018]. Statistical analyses: `lme4`[@bates_fitting_2015], `mgcv`[@wood_generalized_2017]. Data visualisation: `cowplot`[@cowplot_2020], `egg`[@egg_2019], `ggplot2`[@wickham_ggplot2_2016],`ggnewscale`[@ggnewscale_2021], `ggtext`[@ggtext_2020], `gridExtra`[@gridextra_2017],`png`[@png_2013], `RColorBrewer`[@rcolorbrewer_2014],  `rphylopic`[@gearty_rphylopic_2023], `scales`[@scales_2020]. Document preparation: `kableExtra`[@kableextra_2020], `knitr`[@knitr_2015], `rmarkdown`[@rmarkdown_2018].

__3309/3000 words__

# End Notes

## Supplementary Online Material

### Supplementary Figures

__Figure S1__ Schematic diagram of the protocol for defining conservation interventions as in place or not.

<!-- (ref:cap-1) Schematic diagram of the protocol for defining conservation interventions as in place or not. -->

__Figure S2__ The proportion of species documented as threatened by habitat loss and occurring in PAs, with and without use of the Red List data to determine if species occur in PAs.

<!-- (ref:cap-2) The proportion of species documented as threatened by habitat loss and occurring in PAs. Panel (a) represents the results from the main analyses (Figure 1), where the Species Protection Score from Map of Life[@jetz_include_2022] is used to determine whether species occur in PAs (Figure S1a).  For comparison, panel (b) uses an independent but analogous process of overlapping species' Area of Habitat with PAs, and comparing this percentage overlap against species-specific representation thresholds to determine which species occur in PAs (see Methods). The same is done in panel (c), but after first giving precedence to whether the Red List identifies the species as occurring in PAs. Different bar colours denote different Red List categories: Vulnerable (VU; yellow), Endangered (EN; orange), and Critically Endangered (CR; red). Small numbers above the bars represent the total number of species included in our analyses, by taxonomic class. -->

__Figure S3__ The probability of species being documented as threatened by habitat loss and occurring in PAs, with and without use of the Red List data to determine if species occur in PAs.

<!-- (ref:cap-3) The model-predicted probability of species being documented as threatened by habitat loss and occurring in PAs. Panel (a) represents the results from the main analyses (Figure 1), where the Species Protection Score from Map of Life[@jetz_include_2022] is used to determine whether species occur in PAs (Figure S1a).  For comparison, panel (b) uses an independent but analogous process of overlapping species' Area of Habitat with PAs, and comparing this percentage overlap against species-specific representation thresholds to determine which species occur in PAs (see Methods). The same is done in panel (c), but after first giving precedence to whether the Red List identifies the species as occurring in PAs. Point colours denote different Red List categories: Vulnerable (VU; yellow), Endangered (EN; orange), and Critically Endangered (CR; red). Small numbers at the top of the panels denote the sample size. Points are model-predicted values with 95% confidence intervals. -->

__Figure S4__ The proportion of threatened species lacking meaningful habitat protection or any of the other six categories of documented conservation interventions, with and without use of the Red List data to determine if species occur in PAs.

<!-- (ref:cap-4) The proportion of threatened species with no documentation of implementation of any of the six categories of conservation intervention (irrespective of threat). Panel (a) represents the results from the main analyses (Figure 1), where the Species Protection Score from Map of Life[@jetz_include_2022] is used to determine whether species occur in PAs (Figure S1a).  For comparison, panel (b) uses an independent but analogous process of overlapping species' Area of Habitat with PAs, and comparing this percentage overlap against species-specific representation thresholds to determine which species occur in PAs (see Methods). The same is done in panel (c), but after first giving precedence to whether the Red List identifies the species as occurring in PAs. Different bar colours denote different Red List categories: Vulnerable (yellow), Endangered (orange), and Critically Endangered (red). Small numbers above the bars represent the total number of species included in our analyses, by taxonomic class. -->

__Figure S5__ The probability that threatened species lack meaningful habitat protection or any of the other six categories of documented conservation interventions, with and without use of the Red List data to determine if species occur in PAs.

<!-- (ref:cap-5)  The model-predicted probability of threatened species receiving any documented conservation intervention. Panel (a) represents the results from the main analyses (Figure 1), where the Species Protection Score from Map of Life[@jetz_include_2022] is used to determine whether species occur in PAs (Figure S1a).  For comparison, panel (b) uses an independent but analogous process of overlapping species' Area of Habitat with PAs, and comparing this percentage overlap against species-specific representation thresholds to determine which species occur in PAs (see Methods). The same is done in panel (c), but after first giving precedence to whether the Red List identifies the species as occurring in PAs. Point colours denote different Red List categories: Vulnerable (VU; yellow), Endangered (EN; orange), and Critically Endangered (CR; red). Small numbers at the top of the panels denote the sample size. Points are model-predicted values with 95% confidence intervals. -->

__Figure S6__ The probability of threatened species receiving the appropriate type of conservation intervention, excluding cases where the scope or severity of the threat is unknown, or excluding species that are listed as "Possibly Extinct".

<!-- (ref:cap-6) The model-predicted probability of threatened species receiving the appropriate conservation intervention. Results are shown for the three major threats to biodiversity with a clearly matched intervention (rows). Column (a) depicts results from the main analyses. Column (b) depicts results when excluding cases where the scope or severity of the threat is unknown. Column (c) depicts results when excluding species that are listed as "Possibly Extinct". Point colours denote different Red List categories: Vulnerable (yellow), Endangered (orange), and Critically Endangered (red). Small numbers at the top of the panels denote the sample size. Points are model-predicted values with 95% confidence intervals. -->

__Figure S7__ The probability that threatened species lack meaningful habitat protection or any of the other six categories of documented conservation interventions, excluding cases where the scope or severity of the threat is unknown, or excluding species that are listed as "Possibly Extinct".

<!-- (ref:cap-7) The model-predicted probability of threatened species receiving any documented conservation intervention. Column (a) depicts results from the main analyses. Column (b) depicts results when excluding cases where the scope or severity of the threat is unknown. Column (c) depicts results when excluding species that are listed as "Possibly Extinct". Point colours denote different Red List categories: Vulnerable (yellow), Endangered (orange), and Critically Endangered (red). Small numbers at the top of the panels denote the sample size. Points are model-predicted values with 95% confidence intervals. -->

__Figure S8__ The total number of threatened species within each country lacking the appropriate type of conservation intervention.

<!-- (ref:cap-8) The total number of threatened species within each country lacking the appropriate conservation intervention. Results are summarised for each taxonomic group (rows), and for each of the three major threats to biodiversity with a clearly matched intervention (columns). -->

__Figure S9__ The total number of threatened species lacking any documented conservation interventions by country.

<!-- (ref:cap-9) The total number of threatened species lacking any documented conservation interventions, in each country and taxonomic group. -->

__Figure S10__ The total number of threatened species lacking any documented conservation interventions by country, summed across taxonomic classes.

<!-- (ref:cap-10)  The total number of threatened species lacking any documented conservation intervention in each country, summed across the three vertebrate classes that have range data and have been comprehensively assessed: amphibians, birds and mammals (as in Figure 2 in the main text). -->

__Figure S11__ The total number of threatened species lacking any documented conservation interventions by country, summed across the three major threats.

<!-- (ref:cap-11)  The total number of threatened species lacking any documented conservation intervention in each country, summed across the three major threats to biodiversity with a clearly matched intervention. As in Figure 2 in the main text, we consider here only the three vertebrate classes that have range data and have been comprehensively assessed: amphibians, birds and mammals. Note that the same species is counted multiple times if they face more than one of the three major threats. -->

__Figure S12__ The number of threatened endemic species per country, including the number and proportion that lack documented conservation interventions.

<!-- (ref:cap-12) A breakdown of threatened endemic (i.e., found only in one country) species by: (a) the total number of threatened endemic species lacking any documented conservation interventions; (b) the total number of threatened endemic species; and (c) the proportion of threatened endemic species within a country that are lacking any documented conservation interventions. -->

__Figure S13__ The number of threatened endemic species per country with and without documented conservation interventions.

<!-- (ref:cap-13) The number of threatened endemic (i.e., found only in one country) species per country with documented conservation interventions versus without documented interventions. Axes have been transformed to a log1p scale to aid visualisation. Point size denotes the number of observations at each location along the x and y axes.  -->

__Figure S14__ The model-predicted probability that endemic species in each country have no documented conservation interventions.

<!-- (ref:cap-14) The model-predicted probability that endemic species in each country have no documented conservation interventions. Countries are shaded by continent (panels) to ease visual interpretation. Errorbars correspond to 95% Confidence Intervals. -->

__Figure S15__ The number of threatened species per country changed Red List status from 2006-2020.

<!-- (ref:cap-15) A breakdown of species that changed Red List status from 2006-2020 by country, depicting: (a) the total number of threatened species in a country that were downlisted; (b) the total number of threatened species in a country that were uplisted; and (c) the difference between the number of threatened species that were downlisted and the number that were uplisted. In panel (c), negative values (yellow to red) indicate that more species were uplisted than were downlisted, while the reverse is true for positive values (green). -->

__Figure S16__ The model-predicted probability that threatened species have no documented conservation interventions with increasing Evolutionary Distinctiveness.

<!-- (ref:cap-16) The model-predicted probability that threatened species have no documented conservation interventions, with increasing Evolutionary Distinctiveness. Panels separate taxonomic classes, and lines are coloured by Red List category: Vulnerable (yellow), Endangered (orange), and Critically Endangered (red). Shaded bands correspond to 95% Confidence Intervals. Evolutionary Distinctiveness is standardised within taxonomic classes to range between 0 and 1. -->


### Supplementary Tables

__Table S1__ List of taxa comprehensively assessed on the IUCN Red List and included in our analyses.

__Table S2__ Reclassification of IUCN threat categories into simplified categories based on the five main threats recognised by the Convention on Biological Diversity.

__Table S3__ Conservation legislation considered when determining if a species is subject to international legislation.

__Table S4__ Summary of species lacking appropriate type of conservation intervention.

__Table S5__ The number of species changing Red List status from 2006-2020.

__Table S6__ Summary of model results for the GLM of proportion of threatened species with the appropriate type of conservation intervention in place, as a function of taxonomic class and Red List category.

__Table S7__ Summary of model results for the GLM of proportion of threatened species with no documented conservation interventions, as a function of taxonomic class and Red List category.

__Table S8__ Summary of model results for the GLM of proportion of endemic threatened species with no documented conservation interventions, as a function of country.

__Table S9__ Summary of model results for the GLM of model-predicted probability of endemic species receiving no conservation interventions, as a function of the total number of endemic threatened species and country GDP.

### Supplementary Data Files 

```{r sort-supp-dat, eval = FALSE}
data_s1 <- alliucn %>% 
  dplyr::select("scientificName","phylumName","className","familyName","iucn",
                "realms","countries","islandSp_95","ED",
                "threat_habitat_change", "threat_resource_use", "threat_inv_spp",
                "threat_pollution","threat_climate_change","threat_other",
                "inPA_SPS","trade_ctrl","invasive_ctrl_success",
                "education","reintroduction","legislation",
                "matched_hc_sps","matched_ru","matched_is","actions_zero_sps") %>% 
  rename("redList" = "iucn",
         "threat_overexploit" = "threat_resource_use",
         "matched_oe" = "matched_ru")

data_s2 <- statchange %>% 
  dplyr::select("scientificName","phylumName","className","familyName","iucn",
                "cat_old_cat", "cat_change_bi",
                "realms","countries","islandSp_95","ED",
                "threat_habitat_change", "threat_resource_use", "threat_inv_spp",
                "threat_pollution","threat_climate_change","threat_other",
                "inPA_SPS","trade_ctrl","invasive_ctrl_success",
                "education","reintroduction","legislation",
                "matched_hc_sps","matched_ru","matched_is","actions_zero_sps") %>% 
  rename("redList" = "iucn",
         "threat_overexploit" = "threat_resource_use",
         "matched_oe" = "matched_ru",
         "redList_old" = "cat_old_cat", 
         "redList_change" = "cat_change_bi")

write.csv(data_s1, "submission/Nature/R4/datafile_S1.csv", row.names = FALSE)
write.csv(data_s2, "submission/Nature/R4/datafile_S2.csv", row.names = FALSE)
```

__Data S1__ Data for all `r nrow(alliucn)` threatened species including in our analyses of conservation interventions

__Data S2__ Data for all `r nrow(statchange)` threatened species including in our analyses of changes in Red List category

__Data S3__ Metadata for Data S1 and S2

## Data availability 

Processed data to support the findings presented here are available as supplementary data files S1-S3. The original datasets are available for download by request from their respective providers.

### Species data 

* Species assessments can be accessed from the IUCN Red List of Threatened Species website at https://www.iucnredlist.org
* Status change tables are available in pdf format here (Table 7): https://www.iucnredlist.org/resources/summary-statistics
* Range maps can be requested at https://www.iucnredlist.org/resources/spatial-data-download
* Species Protection Scores are available from Map of Life (MOL) at https://mol.org/species/
* Elevation preferences for birds[@quintero_global_2018] available at https://doi.org/10.1038/nature25794
* Point occurrence records for species with zero HSR were downloaded from the Global Biodiversity Information Facility (GBIF)[@gbif_2021] at https://doi.org/10.15468/DL.DVP728
* EDGE data are available at https://www.edgeofexistence.org/edge-lists/

<!-- ### Trait data -->

<!-- * Bird traits: -->
<!--     * EltonTraits 1.0, available at https://doi.org/10.1890/13-1917.1 -->
<!--     * elevation preferences from https://doi.org/10.1038/nature25794 -->
<!-- * Mammal traits = EltonTraits 1.0, available at https://doi.org/10.1890/13-1917.1 -->
<!-- * Amphibian traits = AmphiBIO, available at https://doi.org/10.1038/sdata.2017.123 -->
<!-- * Reptile traits = body mass was calculated from snout-vent length using data and allometric equations from https://doi.org/10.1111/geb.12773 -->

### Conservation interventions data

* Protected Area boundaries can be requested from the World Database on Protected Areas at https://www.unep-wcmc.org/resources-and-data/analysis/main/wdpa
* The Database of Island Invasive Species Eradications[@spatz_globally_2017] is found at http://diise.islandconservation.org/
* International trade control data (CITES, CMS and EU Annexes) is available from Species+ at https://speciesplus.net/

### Geographic data

* Global Administrative Areas available at https://gadm.org/download_country.html 
* Global Islands Database[@sayre_new_2019] available at https://resources.unep-wcmc.org/products/f98e179ec3f448e59dfe9bda248ff4b6 
* Elevation was derived from the EarthEnv Digital Elevation Model Version 1[@robinson_earthenv-dem90:_2014], available at https://www.earthenv.org/DEM
* The global terrestrial habitat types map[@jung_global_2020] is available at https://zenodo.org/records/4058819
* Country GDP is available at https://data.worldbank.org/indicator/NY.GDP.MKTP.CD?locations=1W

## Code availability

Custom Python code to calculate species Area of Habitat in Google Earth Engine is available on request from R.A.S.

## Acknowledgements

We are indebted to David Dorini for help in manually validating the conservation interventions in place for threatened species. We thank IUCN, BirdLife International and UNEP-WCMC for collating and maintaining the public databases used in this study. We thank Dr Stuart Butchart for providing information about the Red List assessment process. Thanks to the three anonymous reviewers, whose feedback greatly improved this study. Funding for this research was provided by the High Meadows Foundation.

## Author contributions

R.A.S. and D.S.W. conceived the study. R.A.S., R.B., D.L., A.K.K., and W.J. collated and processed the data and W.J. contributed guidance on analysis. R.A.S. carried out the analyses and wrote the first draft of the manuscript, with substantial contribution from D.S.W. and R.B. All authors contributed to writing and editing of the manuscript. 

## Competing interests

The authors declare no competing interests.

# References
